Covariate analysis of CMC RNASeq data reprocessed with STAR-feature counts workflow with Ensembl v75 (ACC and DLPFC with CQN normlaisation)

### 2017-07-31 code/Rmd/CMC_ACC_DLPFC_STAR_ENSEMBL_v75_CQN_PENN.Rmd
Minor bug fix

# Add in RIN^2 values
COVARIATES$RIN2 = COVARIATES$RIN^2
ContCovariates = c(ContCovariates, 'RIN2')
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=10, fig.height=8}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
tmp = COVARIATES.CORRELATION$ESTIMATE
tmp[COVARIATES.CORRELATION$PVAL > 0.05] = 0
h = Heatmap(tmp, col = colorRamp2(c(-1,0,1), c('blue','white','red')), name = 'AssocEstimate')
draw(h, heatmap_legend_side = 'left')
```

@@

### Differential expression analysis (SCZ - Control)
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=15, fig.width=20}
# Fit contrast
contrast = makeContrasts(contrasts=c("Dx.TissueSCZ_ACC-Dx.TissueControl_ACC",
                                     "Dx.TissueSCZ_ACC-Dx.TissueOther_ACC",
                                     "Dx.TissueOther_ACC-Dx.TissueControl_ACC",

                                     "Dx.TissueSCZ_DLPFC-Dx.TissueControl_DLPFC",
                                     "Dx.TissueSCZ_DLPFC-Dx.TissueOther_DLPFC",
                                     "Dx.TissueOther_DLPFC-Dx.TissueControl_DLPFC",

                                     "Dx.TissueSCZ_DLPFC-Dx.TissueSCZ_ACC",
                                     "Dx.TissueControl_DLPFC-Dx.TissueControl_ACC",
                                     "Dx.TissueOther_DLPFC-Dx.TissueOther_ACC",

@@

# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = 50000, confint = TRUE) %>%
    rownameToFirstColumn('ensembl_gene_id') 
}, FIT.CONTR) 
names(DE) = gsub('Dx.Tissue', '', colnames(contrast))
names(DE)[10:12] = c('SCZ_ALL-Control_ALL', 'ALL_DLPFC-ALL_ACC', 'ALL_ALL-ALL_ALL')

@@

### Differential expression analysis (ACC - DLPFC, SCZ - Control, Female - Male)
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp2, fig.height=16, fig.width=8}
COVARIATES = tidyr::separate(COVARIATES, Dx.Tissue, c('Dx', 'Tissue')) %>%
  dplyr::mutate(Dx = factor(Dx, levels = c('SCZ', 'Control', 'Other')),
                Tissue = factor(Tissue))
postAdjustCovars = c('Tissue', 'Dx', 'Institution', 'Gender', 'RIN', 'IntronicRate')

# Post adjusted design matrix
DESIGN2 = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F],Intercept = F)$design
DESIGN2 = DESIGN2[,linColumnFinder(DESIGN2)$indepCols]

@@

# Fit contrast
contrast = makeContrasts(contrasts=c("TissueACC-TissueDLPFC",
                                     "DxSCZ-DxControl",
                                     "GenderFemale"),
                         levels = colnames(ADJUSTED.FIT2$coefficients))
FIT.CONTR2 = contrasts.fit(ADJUSTED.FIT2, contrasts=contrast)
FIT.CONTR2 = eBayes(FIT.CONTR2)

>>>>

# Add in RIN^2 values
COVARIATES$RIN2 = COVARIATES$RIN^2
ContCovariates = c(ContCovariates, 'RIN2')
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=10, fig.height=8}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES[,setdiff(colnames(COVARIATES), 'Institution')], PVAL = 0.05)
tmp = COVARIATES.CORRELATION$ESTIMATE
tmp[COVARIATES.CORRELATION$PVAL > 0.05] = 0
h = Heatmap(tmp, col = colorRamp2(c(-1,0,1), c('blue','white','red')), name = 'AssocEstimate')
draw(h, heatmap_legend_side = 'left')
```

@@

### Differential expression analysis (SCZ - Control)
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=15, fig.width=20}
# Fit contrast
contrast = makeContrasts(contrasts=c("Dx.TissueSCZ_ACC-Dx.TissueControl_ACC",

                                     "Dx.TissueSCZ_DLPFC-Dx.TissueControl_DLPFC",

                                     "Dx.TissueSCZ_DLPFC-Dx.TissueSCZ_ACC",
                                     "Dx.TissueControl_DLPFC-Dx.TissueControl_ACC",

@@

# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = 50000, confint = TRUE) %>%
    rownameToFirstColumn('ensembl_gene_id') 
}, FIT.CONTR) 
names(DE) = gsub('Dx.Tissue', '', colnames(contrast))
names(DE)[5:7] = c('SCZ_ALL-Control_ALL', 'ALL_DLPFC-ALL_ACC', 'ALL_ALL-ALL_ALL')

@@

### Differential expression analysis (ACC - DLPFC, SCZ - Control, Female - Male)
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp2, fig.height=16, fig.width=8}
COVARIATES = tidyr::separate(COVARIATES, Dx.Tissue, c('Dx', 'Tissue')) %>%
  dplyr::mutate(Dx = factor(Dx, levels = c('SCZ', 'Control', 'Other')),
                Tissue = factor(Tissue))
postAdjustCovars = c('Tissue', 'Dx', 'Gender', 'RIN', 'IntronicRate')

# Post adjusted design matrix
DESIGN2 = getDesignMatrix(droplevels(COVARIATES[,postAdjustCovars,drop=F]),Intercept = T)$design
DESIGN2 = DESIGN2[,linColumnFinder(DESIGN2)$indepCols]

@@

# Fit contrast
contrast = makeContrasts(contrasts=c("TissueDLPFC",
                                     "DxControl",
                                     "GenderMale"),
                         levels = colnames(ADJUSTED.FIT2$coefficients))
FIT.CONTR2 = contrasts.fit(ADJUSTED.FIT2, contrasts=contrast)
FIT.CONTR2 = eBayes(FIT.CONTR2)
